---
title: "Covid91 vaccine study Final2020"
author: "Bethanee Olmos"
date: "October 2020"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

In this clinical trial, is the vaccine effective in preventing the targeted population from getting COVID 19?

$H_0$: The null hypothesis is that among the target population, the vaccine has no effect on whether one contracts the virus or not.

$H_A$: The alternative hypothesis is that among the target population, the vaccine has an effect on whether one contracts the virus or not.

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

Every question is of the categorical~categorical type.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



##Graphical Descriptive Results

# Female Report
  

```{r}
DataF = filter(FinalData ,sex=="gal") 
barchartGC(~infected + treatment,data=DataF)
barchartGC(~infected + treatment,data=DataF, type="percent", main="gal")

```


Among the females who contracted covid-19 twice as many took the placebo. Therefore it appears that the vaccine does have a positive effect. 


# Male Report

```{r}
DataG = filter(FinalData,sex=="guy") 
barchartGC(~infected + treatment,data=DataG)
barchartGC(~infected + treatment,data=DataG, type="percent", main="guys")

```

From the graph above, it appears that the vaccine has a negative effect on guys since there is more than a 55% chance for them to get COVID if they receive the vaccine.


#LGBTQ Report

```{r}
DataLGBTQ = filter (FinalData, LGBTQ=="gay")
barchartGC(~infected + treatment, data= DataLGBTQ)
barchartGC(~infected + treatment, data=DataLGBTQ, type="percent", main="gays")
```


From the graph above, it appears that the vaccine has a negative effect on gay population since there is more than 60 % of chance for them to get COVID if they receive the vaccine.
  
# Druguser report  

  

```{r}
DataDrugUser = filter (FinalData, DrugUser=="yes")
barchartGC(~infected + treatment, data= DataDrugUser)
barchartGC(~infected + treatment, data= DataDrugUser, type="percent", main="DrugUser")
```
  
  
It appears in the chart above that the vaccine had a positive effect on drug users because less than 40% of the population who received the vaccine got covid19.


##Numerical Inferential Results

```{r}
DataF = filter(FinalData,sex=="gal")
table1 <- xtabs(~infected + treatment, data=DataF)
rowPerc(table1)
colPerc(table1)
```

According to the above table, only 4.44% of females who received the vaccine contracted COVID, while 95.56% did not. 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value is 3.626e-16 < 0.05 so the possibility of the vaccine being more effective for females than the placebo is extremely unlikely.

```{r}
DataG = filter(FinalData,sex=="guy")
table1 <- xtabs(~infected + treatment, data=DataG)
rowPerc(table1)
colPerc(table1)
```

According to the above table, only 3.72% of guys who received the vaccine contracted COVID, while 96.28% did not. 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value is 0.0006181 < 0.05 so the possibility of the vaccine being more effective for males than the placebo is extremely unlikely.


```{r}
DataLGBTQ = filter(FinalData,LGBTQ=="gay")
table1 <- xtabs(~infected + treatment, data=DataLGBTQ)
rowPerc(table1)
colPerc(table1)
```

According to the above table, only 7.97% of the LGBTQ population who received the vaccine contracted COVID, while 92.03% did not.   


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p value is 1.05e-05 < 0.05 so the possibility of the vaccine being more effective for the LGBTQ population than the placebo is extremely unlikely.


```{r}
DataDrugUser = filter(FinalData,DrugUser=="yes")
table1 <- xtabs(~infected + treatment, data=DataDrugUser)
rowPerc(table1)
colPerc(table1)
```

According to the above table, only 3.63% of Drug users who received the vaccine contracted COVID, while 96.37% did not. 


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
The p value is 0.003028 < 0.05 so the possibility of the vaccine being more effective for the Drug user population than the placebo is extremely unlikely. 
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->